Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-02-13 vertical-thermal-review.Rmd

## 3.2. Leaf optical properties
<!---need to cut parts of this paragraph or merge into 3.1?-->
Leaf optical properties of leaves are influenced by a combination of anatomical, morphological, and biochemical traits that vary vertically, as outlined above.
Along the vertical light gradient, leaf light absorptance is balanced with $T_{leaf}$ regulation, where excess light can impair photosynthetic machinery (see below), and is only modestly greater in sun leaves or mid-canopy leaves than shade leaves. 
In higher light environment, PAR absorptance and thus heat load is reduced by reflecting light through leaf surface modifications as mentioned above, such as trichome and cuticle layer, or through biochemically dissipating heat through caretenoid concentrations [Table 1, @knappVariabilityLeafOptical1998; @leeCorrelatesLeafOptical1990].
Further, there is less need for light absorption efficiency, but greater need to channel the higher availability of light into deeper cell layers through "light pipes" or palisade layers that increase path length for light penetration, contributing to higher photosynthetic capacity.
The thickness of sun leaves also reduces light transmittance or light scattered after entering the leaf more than in shade leaves [@poorterLeafOpticalProperties1995; @leeCorrelatesLeafOptical1990].
However, in the understory or shade regions where PAR is more limiting, leaf light absorption efficiency is greater, enhanced by shade leaves maintaining higher SLA, denser layers of chloroplasts, shorter, fewer layers of conical palisade cells, higher nitrogen, and chlorophyll per unit biomass that exemplifies optical absorption capabilities, resulting in similar chlorophyll concentration per unit area as in sun leaves. 
Therefore, shade leaves can have comparable light harvesting capabilities as in sun leaves, and higher absorption efficiency per unit biomass [@poorterLeafOpticalProperties1995; @poorterLeafOpticalProperties2000; @leeCorrelatesLeafOptical1990]. 
Thus, while canopy leaves tend to invest more energy in metabolic activity and heat dissipation, understory leaves invest in light capture.

>>>>

## 3.2. Leaf optical properties
<!---need to cut parts of this paragraph or merge into 3.1?-->
Leaf optical properties of leaves are influenced by a combination of anatomical, morphological, and biochemical traits that vary vertically, as outlined above.
Along the vertical light gradient, leaf light absorptance is balanced with $T_{leaf}$ regulation, where excess light can impair photosynthetic machinery (see below), and is only modestly greater in sun leaves or mid-canopy leaves than shade leaves. 
In higher light environment, PAR absorptance and thus heat load is reduced by reflecting light through leaf surface modifications as mentioned above, such as trichome and cuticle layer, or through biochemically dissipating heat through caretenoid concentrations [Table `r table_leaf_traits`, @knappVariabilityLeafOptical1998; @leeCorrelatesLeafOptical1990].
Further, there is less need for light absorption efficiency, but greater need to channel the higher availability of light into deeper cell layers through "light pipes" or palisade layers that increase path length for light penetration, contributing to higher photosynthetic capacity.
The thickness of sun leaves also reduces light transmittance or light scattered after entering the leaf more than in shade leaves [@poorterLeafOpticalProperties1995; @leeCorrelatesLeafOptical1990].
However, in the understory or shade regions where PAR is more limiting, leaf light absorption efficiency is greater, enhanced by shade leaves maintaining higher SLA, denser layers of chloroplasts, shorter, fewer layers of conical palisade cells, higher nitrogen, and chlorophyll per unit biomass that exemplifies optical absorption capabilities, resulting in similar chlorophyll concentration per unit area as in sun leaves. 
Therefore, shade leaves can have comparable light harvesting capabilities as in sun leaves, and higher absorption efficiency per unit biomass [@poorterLeafOpticalProperties1995; @poorterLeafOpticalProperties2000; @leeCorrelatesLeafOptical1990]. 
Thus, while canopy leaves tend to invest more energy in metabolic activity and heat dissipation, understory leaves invest in light capture.
